Appendix S1: Supplemental figures and table

### 2024-01-22 ms/si.qmd
copyediting edits for si

---
title: "Appendix S1: Supplemental figures and table"
format:
  pdf:
    keep-tex: true  
    include-in-header: 
      - preamble.tex
    toc: false
header-includes: |
  \usepackage{fancyhdr}
  \pagestyle{fancy}
  \fancyhead[C]{Triplett \textit{et al.} — \textit{American Journal of Botany} 2024 – Appendix S1}
  \addtokomafont{disposition}{\rmfamily}
  \RedeclareSectionCommand[
    font=\normalfont\Large]{section}
  \RedeclareSectionCommand[
    font=\normalfont\normalsize\bfseries]{subsection}
  \RedeclareSectionCommand[
    font=\normalfont\normalsize\itshape]{subsubsection}
  \RedeclareSectionCommand[
    font=\normalfont\normalsize]{paragraph}
mainfont: Times New Roman
---

@@

```{r, sum-data, echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
#| label: sum-data
#| tbl-cap: "Average traits values among ʻilima leaves at each site. $\\mathrm{SD}_\\text{abaxial}$ is the stomatal density per mm$^2$ on the abaxial (lower) surface; $\\mathrm{SD}_\\text{adaxial}$ is the stomatal density per mm$^2$ on the adaxial (upper) surface; $\\mathrm{GCL}_\\text{abaxial}$ is the guard cell length in $\\mu$m on the abaxial (lower) surface; $\\mathrm{GCL}_\\text{adaxial}$ is the guard cell length in $\\mu$m on the adaxial (upper) surface; Leaf thickness is the length from upper cuticle to lower cuticle in $\\mu$m; $A$ is the photosynthetic rate in $\\mu \\text{mol CO}_2~\\text{m}^{-2}~\\text{s}^{-1}$; $g_\\mathrm{sw}$ is the stomatal conductance to water vapor in $\\text{mol}~\\text{m}^{-2}~\\text{s}^{-1}$."
#| 
sum_data |>
  knitr::kable(format = "latex", digits = c(rep(0, 3), rep(2, 6), 3),
               booktabs = TRUE, escape = FALSE, longtable = TRUE) |>
  kableExtra::kable_styling(latex_options = c("striped", "no_vertical_lines")) |>
  kableExtra::column_spec(1, width = "4cm") |>
  kableExtra::column_spec(2, width = "1.5cm") |>
  kableExtra::column_spec(3, width = "1.5cm") |>
  kableExtra::landscape()

@@

```{=tex}
\begin{figure}
  \includegraphics{../figures/ags-curve.pdf}
  \caption{Idealized method for collecting $A \textendash g_\text{sw}$ curves on either amphi or pseudohypo leaves. A) After clamping the leaf into the LI-6800 chamber, it acclimates to high light ($\mathrm{PPFD} = 2000~\mu \text{mol}~\text{m}^{-2}~\text{s}^{-1}$) and humidity ($\mathrm{RH} = 70\%$). This induces stomata to open, increasing $g_\mathrm{sw}$ and $A$ until they reach a maximum. We abruptly lower the chamber humidity to $\approx 10\%$ to close stomata and log data (black points) until $g_\mathrm{sw}$ and $A$ reach their nadir. B) We fit $A \textendash g_\text{sw}$ curves to logged data points. The asterisk in both panels indicates the data point used for maximum $A$ and $g_\mathrm{sw}$. C) $\mathrm{AA}$ is low (left panel) when the photosynthetic rate of an amphi leaf is similar to a pseudohypo leaf at the same total $g_\mathrm{sw}$ ($x$-axes); large $\mathrm{AA}$ (right panel) is indicated when an amphi leaf has a higher photosynthetic rate than a pseudohypo leaf. Abbreviations: $A$ is the photosynthetic rate; $\mathrm{AA}$ is the amphistomy advantage; $g_\mathrm{sw}$ is the stomatal conductance to water vapor; $\mathrm{PPFD}$ is photosynthetic photon flux density; $\mathrm{RH}$ is relative humidity.}
  \label{fig:ags-curve}
\end{figure}
```

@@

```{=tex}
\begin{figure}
  \caption{(Continued from previous page.) Individual $A \textendash g_\mathrm{sw}$ curves used to estimate $\mathrm{AA}$. For each leaf, one per site, we measured $A$ ($y$-axis) over a range of $g_\mathrm{sw}$ ($x$-axis) on the same leaf with two treatments: `amphi' (circles, solid line) leaves were untreated; `pseudohypo' (squares, dashed line) leaves had no conductance through the upper (adaxial) surface. In all coastal (orange) and montane (blue) leaves, we fit generalized additive models and 95\% confidence ribbons to estimate $\mathrm{AA}$ at a $g_\mathrm{sw}$ where the curves overlap (vertical black line). In leaves from Kaloko-Honokōhau national historical park and Puakō petroglyph park, we extrapolated sligtly beyond fitted curves because they did not quite overlap. Symbols: $\mathrm{AA}$, amphistomy advantage; $A$, photosynthetic rate in $\mu \text{mol CO}_2~\text{m}^{-2}~\text{s}^{-1}$; $g_\mathrm{sw}$, stomatal conductance to water vapor in $\text{mol}~\text{m}^{-2}~\text{s}^{-1}$.}
  \label{fig:licor}
\end{figure}
```

@@

```{=tex}
\begin{figure}
  \includegraphics{../figures/pp-licor.pdf}
  \caption{Posterior predictions (thin grey lines) from fitted $A \textendash g_\mathrm{sw}$ curves closely match the observed distribution (thick black line), indicating that the statistical model adequately captures variation in the response variable over the measured range. Symbols: $A$, photosynthetic rate in $\mu \text{mol CO}_2~\text{m}^{-2}~\text{s}^{-1}$; $g_\mathrm{sw}$, stomatal conductance to water vapor in $\text{mol}~\text{m}^{-2}~\text{s}^{-1}$.}
  \label{fig:pp-licor}
\end{figure}
```

@@

```{=tex}
\begin{figure}
  \includegraphics{../figures/habitat-Ags.pdf}
  \caption{The photsynthetic rate (left facet) and stomatal conductance to water vapor (right facet) of montane (blue) and coastal (orange) ʻilima leaves. Each point-interval is the median posterior estimate plus 95\% confidence interval of trait value for that habitat. Smaller points next to each point-interval are the $g_\mathrm{smax,ratio}$ of individual plants, one per site. Symbols: $A$, photosynthetic rate in $\mu \text{mol CO}_2~\text{m}^{-2}~\text{s}^{-1}$; $g_\mathrm{sw}$, stomatal conductance to water vapor in $\text{mol}~\text{m}^{-2}~\text{s}^{-1}$.}
  \label{fig:habitat-Ags}
\end{figure}
```

>>>>

---
title: "Appendix S1: Supplemental figures and table"
format:
  pdf:
    keep-tex: true  
    include-in-header: 
      - preamble.tex
    toc: false
header-includes: |
  \usepackage{textgreek}
  \usepackage{fancyhdr}
  \pagestyle{fancy}
  \fancyhead[C]{Triplett \textit{et al.} — \textit{American Journal of Botany} 2024 – Appendix S1}
  \addtokomafont{disposition}{\rmfamily}
  \RedeclareSectionCommand[
    font=\normalfont\Large]{section}
  \RedeclareSectionCommand[
    font=\normalfont\normalsize\bfseries]{subsection}
  \RedeclareSectionCommand[
    font=\normalfont\normalsize\itshape]{subsubsection}
  \RedeclareSectionCommand[
    font=\normalfont\normalsize]{paragraph}
mainfont: Times New Roman
---

@@

```{r, sum-data, echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
#| label: sum-data
#| tbl-cap: "Average traits values among ʻilima leaves at each site. $\\mathrm{SD}_\\text{abaxial}$ is the stomatal density per mm$^2$ on the abaxial (lower) surface; $\\mathrm{SD}_\\text{adaxial}$ is the stomatal density per mm$^2$ on the adaxial (upper) surface; $\\mathrm{GCL}_\\text{abaxial}$ is the guard cell length in \\textmu m on the abaxial (lower) surface; $\\mathrm{GCL}_\\text{adaxial}$ is the guard cell length in \\textmu m on the adaxial (upper) surface; Leaf thickness is the length from upper cuticle to lower cuticle in \\textmu m; $A$ is the photosynthetic rate in \\textmu mol CO$_2~\\text{m}^{-2}~\\text{s}^{-1}$; $g_\\mathrm{sw}$ is the stomatal conductance to water vapor in $\\text{mol}~\\text{m}^{-2}~\\text{s}^{-1}$."
#| 
sum_data |>
  knitr::kable(format = "latex", digits = c(rep(0, 3), rep(2, 6), 3),
               booktabs = TRUE, escape = FALSE, longtable = TRUE) |>
  kableExtra::kable_styling(latex_options = c("striped", "no_vertical_lines")) |>
  kableExtra::column_spec(1, width = "4cm") |>
  kableExtra::column_spec(2, width = "1.5cm") |>
  kableExtra::column_spec(3, width = "1.5cm") |>
  kableExtra::landscape()

@@

```{=tex}
\begin{figure}
  \includegraphics{../figures/ags-curve.pdf}
  \caption{Idealized method for collecting $A \textendash g_\text{sw}$ curves on either amphi or pseudohypo leaves. (A) After clamping the leaf into the LI-6800 chamber, it acclimates to high light ($\mathrm{PPFD} = 2000~\textmu \text{mol}~\text{m}^{-2}~\text{s}^{-1}$) and humidity ($\mathrm{RH} = 70\%$). This induces stomata to open, increasing $g_\mathrm{sw}$ and $A$ until they reach a maximum. We abruptly lower the chamber humidity to $\approx 10\%$ to close stomata and log data (black points) until $g_\mathrm{sw}$ and $A$ reach their nadir. (B) We fit $A \textendash g_\text{sw}$ curves to logged data points. The asterisk in both panels indicates the data point used for maximum $A$ and $g_\mathrm{sw}$. (C) $\mathrm{AA}$ is low (left panel) when the photosynthetic rate of an amphi leaf is similar to a pseudohypo leaf at the same total $g_\mathrm{sw}$ ($x$-axes); large $\mathrm{AA}$ (right panel) is indicated when an amphi leaf has a higher photosynthetic rate than a pseudohypo leaf. Abbreviations: $A$ is the photosynthetic rate; $\mathrm{AA}$ is the amphistomy advantage; $g_\mathrm{sw}$ is the stomatal conductance to water vapor; $\mathrm{PPFD}$ is photosynthetic photon flux density; $\mathrm{RH}$ is relative humidity.}
  \label{fig:ags-curve}
\end{figure}
```

@@

```{=tex}
\begin{figure}
  \caption{(Continued from previous page.) Individual $A \textendash g_\mathrm{sw}$ curves used to estimate $\mathrm{AA}$. For each leaf, one per site, we measured $A$ ($y$-axis) over a range of $g_\mathrm{sw}$ ($x$-axis) on the same leaf with two treatments: `amphi' (circles, solid line) leaves were untreated; `pseudohypo' (squares, dashed line) leaves had no conductance through the upper (adaxial) surface. In all coastal (orange) and montane (blue) leaves, we fit generalized additive models and 95\% confidence ribbons to estimate $\mathrm{AA}$ at a $g_\mathrm{sw}$ where the curves overlap (vertical black line). In leaves from Kaloko-Honokōhau national historical park and Puakō petroglyph park, we extrapolated sligtly beyond fitted curves because they did not quite overlap. Symbols: $\mathrm{AA}$, amphistomy advantage; $A$, photosynthetic rate in \textmu mol CO$_2~\text{m}^{-2}~\text{s}^{-1}$; $g_\mathrm{sw}$, stomatal conductance to water vapor in $\text{mol}~\text{m}^{-2}~\text{s}^{-1}$.}
  \label{fig:licor}
\end{figure}
```

@@

```{=tex}
\begin{figure}
  \includegraphics{../figures/pp-licor.pdf}
  \caption{Posterior predictions (thin grey lines) from fitted $A \textendash g_\mathrm{sw}$ curves closely match the observed distribution (thick black line), indicating that the statistical model adequately captures variation in the response variable over the measured range. Symbols: $A$, photosynthetic rate in \textmu mol CO$_2~\text{m}^{-2}~\text{s}^{-1}$; $g_\mathrm{sw}$, stomatal conductance to water vapor in $\text{mol}~\text{m}^{-2}~\text{s}^{-1}$.}
  \label{fig:pp-licor}
\end{figure}
```

@@

```{=tex}
\begin{figure}
  \includegraphics{../figures/habitat-Ags.pdf}
  \caption{The photsynthetic rate (left panel) and stomatal conductance to water vapor (right panel) of montane (blue) and coastal (orange) ʻilima leaves. Each point-interval is the median posterior estimate plus 95\% confidence interval of trait value for that habitat. Smaller points next to each point-interval are the $g_\mathrm{smax,ratio}$ of individual plants, one per site. Symbols: $A$, photosynthetic rate in \textmu mol CO$_2~\text{m}^{-2}~\text{s}^{-1}$; $g_\mathrm{sw}$, stomatal conductance to water vapor in $\text{mol}~\text{m}^{-2}~\text{s}^{-1}$.}
  \label{fig:habitat-Ags}
\end{figure}
```

### 2024-01-22 ms/si.tex
copyediting edits for si

% make better hyphen for numeric ranges, \numrange{}{}
\usepackage[detect-none]{siunitx}
\sisetup{range-phrase = \text{--}}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\KOMAoption{captions}{tableheading}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhead[C]{Triplett \textit{et al.} — \textit{American Journal of Botany} 2024 – Appendix S1}
\addtokomafont{disposition}{\rmfamily}
\RedeclareSectionCommand[
  font=\normalfont\Large]{section}
\RedeclareSectionCommand[
  font=\normalfont\normalsize\bfseries]{subsection}
\RedeclareSectionCommand[
  font=\normalfont\normalsize\itshape]{subsubsection}
\RedeclareSectionCommand[
  font=\normalfont\normalsize]{paragraph}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Appendix S1: Supplemental figures and table},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, breakable, boxrule=0pt, enhanced, frame hidden]}{\end{tcolorbox}}\fi

@@

\begin{landscape}
\begin{longtable}{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{1.5cm}>{\raggedright\arraybackslash}p{1.5cm}rrrrrrr}
\caption{Average traits values among ʻilima leaves at each site.
\(\mathrm{SD}_\text{abaxial}\) is the stomatal density per mm\(^2\) on
the abaxial (lower) surface; \(\mathrm{SD}_\text{adaxial}\) is the
stomatal density per mm\(^2\) on the adaxial (upper) surface;
\(\mathrm{GCL}_\text{abaxial}\) is the guard cell length in \(\mu\)m on
the abaxial (lower) surface; \(\mathrm{GCL}_\text{adaxial}\) is the
guard cell length in \(\mu\)m on the adaxial (upper) surface; Leaf
thickness is the length from upper cuticle to lower cuticle in \(\mu\)m;
\(A\) is the photosynthetic rate in
\(\mu \text{mol CO}_2~\text{m}^{-2}~\text{s}^{-1}\); \(g_\mathrm{sw}\)
is the stomatal conductance to water vapor in
\(\text{mol}~\text{m}^{-2}~\text{s}^{-1}\).}\tabularnewline

@@

\begin{figure}
  \includegraphics{../figures/ags-curve.pdf}
  \caption{Idealized method for collecting $A \textendash g_\text{sw}$ curves on either amphi or pseudohypo leaves. A) After clamping the leaf into the LI-6800 chamber, it acclimates to high light ($\mathrm{PPFD} = 2000~\mu \text{mol}~\text{m}^{-2}~\text{s}^{-1}$) and humidity ($\mathrm{RH} = 70\%$). This induces stomata to open, increasing $g_\mathrm{sw}$ and $A$ until they reach a maximum. We abruptly lower the chamber humidity to $\approx 10\%$ to close stomata and log data (black points) until $g_\mathrm{sw}$ and $A$ reach their nadir. B) We fit $A \textendash g_\text{sw}$ curves to logged data points. The asterisk in both panels indicates the data point used for maximum $A$ and $g_\mathrm{sw}$. C) $\mathrm{AA}$ is low (left panel) when the photosynthetic rate of an amphi leaf is similar to a pseudohypo leaf at the same total $g_\mathrm{sw}$ ($x$-axes); large $\mathrm{AA}$ (right panel) is indicated when an amphi leaf has a higher photosynthetic rate than a pseudohypo leaf. Abbreviations: $A$ is the photosynthetic rate; $\mathrm{AA}$ is the amphistomy advantage; $g_\mathrm{sw}$ is the stomatal conductance to water vapor; $\mathrm{PPFD}$ is photosynthetic photon flux density; $\mathrm{RH}$ is relative humidity.}
  \label{fig:ags-curve}
\end{figure}

@@

\begin{figure}
  \caption{(Continued from previous page.) Individual $A \textendash g_\mathrm{sw}$ curves used to estimate $\mathrm{AA}$. For each leaf, one per site, we measured $A$ ($y$-axis) over a range of $g_\mathrm{sw}$ ($x$-axis) on the same leaf with two treatments: `amphi' (circles, solid line) leaves were untreated; `pseudohypo' (squares, dashed line) leaves had no conductance through the upper (adaxial) surface. In all coastal (orange) and montane (blue) leaves, we fit generalized additive models and 95\% confidence ribbons to estimate $\mathrm{AA}$ at a $g_\mathrm{sw}$ where the curves overlap (vertical black line). In leaves from Kaloko-Honokōhau national historical park and Puakō petroglyph park, we extrapolated sligtly beyond fitted curves because they did not quite overlap. Symbols: $\mathrm{AA}$, amphistomy advantage; $A$, photosynthetic rate in $\mu \text{mol CO}_2~\text{m}^{-2}~\text{s}^{-1}$; $g_\mathrm{sw}$, stomatal conductance to water vapor in $\text{mol}~\text{m}^{-2}~\text{s}^{-1}$.}
  \label{fig:licor}
\end{figure}

@@

\begin{figure}
  \includegraphics{../figures/pp-licor.pdf}
  \caption{Posterior predictions (thin grey lines) from fitted $A \textendash g_\mathrm{sw}$ curves closely match the observed distribution (thick black line), indicating that the statistical model adequately captures variation in the response variable over the measured range. Symbols: $A$, photosynthetic rate in $\mu \text{mol CO}_2~\text{m}^{-2}~\text{s}^{-1}$; $g_\mathrm{sw}$, stomatal conductance to water vapor in $\text{mol}~\text{m}^{-2}~\text{s}^{-1}$.}
  \label{fig:pp-licor}
\end{figure}

@@

\begin{figure}
  \includegraphics{../figures/habitat-Ags.pdf}
  \caption{The photsynthetic rate (left facet) and stomatal conductance to water vapor (right facet) of montane (blue) and coastal (orange) ʻilima leaves. Each point-interval is the median posterior estimate plus 95\% confidence interval of trait value for that habitat. Smaller points next to each point-interval are the $g_\mathrm{smax,ratio}$ of individual plants, one per site. Symbols: $A$, photosynthetic rate in $\mu \text{mol CO}_2~\text{m}^{-2}~\text{s}^{-1}$; $g_\mathrm{sw}$, stomatal conductance to water vapor in $\text{mol}~\text{m}^{-2}~\text{s}^{-1}$.}
  \label{fig:habitat-Ags}
\end{figure}

>>>>

% make better hyphen for numeric ranges, \numrange{}{}
\usepackage[detect-none]{siunitx}
\sisetup{range-phrase = \text{--}}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\KOMAoption{captions}{tableheading}
\usepackage{textgreek}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhead[C]{Triplett \textit{et al.} — \textit{American Journal of Botany} 2024 – Appendix S1}
\addtokomafont{disposition}{\rmfamily}
\RedeclareSectionCommand[
  font=\normalfont\Large]{section}
\RedeclareSectionCommand[
  font=\normalfont\normalsize\bfseries]{subsection}
\RedeclareSectionCommand[
  font=\normalfont\normalsize\itshape]{subsubsection}
\RedeclareSectionCommand[
  font=\normalfont\normalsize]{paragraph}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Appendix S1: Supplemental figures and table},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, frame hidden, enhanced, breakable, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

\begin{landscape}
\begin{longtable}{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{1.5cm}>{\raggedright\arraybackslash}p{1.5cm}rrrrrrr}
\caption{Average traits values among ʻilima leaves at each site.
\(\mathrm{SD}_\text{abaxial}\) is the stomatal density per mm\(^2\) on
the abaxial (lower) surface; \(\mathrm{SD}_\text{adaxial}\) is the
stomatal density per mm\(^2\) on the adaxial (upper) surface;
\(\mathrm{GCL}_\text{abaxial}\) is the guard cell length in \textmu m on
the abaxial (lower) surface; \(\mathrm{GCL}_\text{adaxial}\) is the
guard cell length in \textmu m on the adaxial (upper) surface; Leaf
thickness is the length from upper cuticle to lower cuticle in
\textmu m; \(A\) is the photosynthetic rate in \textmu mol
CO\(_2~\text{m}^{-2}~\text{s}^{-1}\); \(g_\mathrm{sw}\) is the stomatal
conductance to water vapor in
\(\text{mol}~\text{m}^{-2}~\text{s}^{-1}\).}\tabularnewline

@@

\begin{figure}
  \includegraphics{../figures/ags-curve.pdf}
  \caption{Idealized method for collecting $A \textendash g_\text{sw}$ curves on either amphi or pseudohypo leaves. (A) After clamping the leaf into the LI-6800 chamber, it acclimates to high light ($\mathrm{PPFD} = 2000~\textmu \text{mol}~\text{m}^{-2}~\text{s}^{-1}$) and humidity ($\mathrm{RH} = 70\%$). This induces stomata to open, increasing $g_\mathrm{sw}$ and $A$ until they reach a maximum. We abruptly lower the chamber humidity to $\approx 10\%$ to close stomata and log data (black points) until $g_\mathrm{sw}$ and $A$ reach their nadir. (B) We fit $A \textendash g_\text{sw}$ curves to logged data points. The asterisk in both panels indicates the data point used for maximum $A$ and $g_\mathrm{sw}$. (C) $\mathrm{AA}$ is low (left panel) when the photosynthetic rate of an amphi leaf is similar to a pseudohypo leaf at the same total $g_\mathrm{sw}$ ($x$-axes); large $\mathrm{AA}$ (right panel) is indicated when an amphi leaf has a higher photosynthetic rate than a pseudohypo leaf. Abbreviations: $A$ is the photosynthetic rate; $\mathrm{AA}$ is the amphistomy advantage; $g_\mathrm{sw}$ is the stomatal conductance to water vapor; $\mathrm{PPFD}$ is photosynthetic photon flux density; $\mathrm{RH}$ is relative humidity.}
  \label{fig:ags-curve}
\end{figure}

@@

\begin{figure}
  \caption{(Continued from previous page.) Individual $A \textendash g_\mathrm{sw}$ curves used to estimate $\mathrm{AA}$. For each leaf, one per site, we measured $A$ ($y$-axis) over a range of $g_\mathrm{sw}$ ($x$-axis) on the same leaf with two treatments: `amphi' (circles, solid line) leaves were untreated; `pseudohypo' (squares, dashed line) leaves had no conductance through the upper (adaxial) surface. In all coastal (orange) and montane (blue) leaves, we fit generalized additive models and 95\% confidence ribbons to estimate $\mathrm{AA}$ at a $g_\mathrm{sw}$ where the curves overlap (vertical black line). In leaves from Kaloko-Honokōhau national historical park and Puakō petroglyph park, we extrapolated sligtly beyond fitted curves because they did not quite overlap. Symbols: $\mathrm{AA}$, amphistomy advantage; $A$, photosynthetic rate in \textmu mol CO$_2~\text{m}^{-2}~\text{s}^{-1}$; $g_\mathrm{sw}$, stomatal conductance to water vapor in $\text{mol}~\text{m}^{-2}~\text{s}^{-1}$.}
  \label{fig:licor}
\end{figure}

@@

\begin{figure}
  \includegraphics{../figures/pp-licor.pdf}
  \caption{Posterior predictions (thin grey lines) from fitted $A \textendash g_\mathrm{sw}$ curves closely match the observed distribution (thick black line), indicating that the statistical model adequately captures variation in the response variable over the measured range. Symbols: $A$, photosynthetic rate in \textmu mol CO$_2~\text{m}^{-2}~\text{s}^{-1}$; $g_\mathrm{sw}$, stomatal conductance to water vapor in $\text{mol}~\text{m}^{-2}~\text{s}^{-1}$.}
  \label{fig:pp-licor}
\end{figure}

@@

\begin{figure}
  \includegraphics{../figures/habitat-Ags.pdf}
  \caption{The photsynthetic rate (left panel) and stomatal conductance to water vapor (right panel) of montane (blue) and coastal (orange) ʻilima leaves. Each point-interval is the median posterior estimate plus 95\% confidence interval of trait value for that habitat. Smaller points next to each point-interval are the $g_\mathrm{smax,ratio}$ of individual plants, one per site. Symbols: $A$, photosynthetic rate in \textmu mol CO$_2~\text{m}^{-2}~\text{s}^{-1}$; $g_\mathrm{sw}$, stomatal conductance to water vapor in $\text{mol}~\text{m}^{-2}~\text{s}^{-1}$.}
  \label{fig:habitat-Ags}
\end{figure}
